---
title: "Assignment of COPD Incidence in Scotland"
author: "by Fahmida Akter"
output:
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load Dataset

```{r load-data}
library(tidyverse)
data <- read_csv("E:/DG Health Assignment/scotpho_data_extract_COPD.csv")

```

## Data Preperation

```{r data-preperation}
#Variable list
names(data)
#head of data
head(data)
#tail of data
tail(data)
#dimension of data (variable and observation)
dim(data)
#checking of missing data
mis <- is.na(data)
sum(mis)

```


## Data Visualization

# Histogram

```{r plot-hist, echo=TRUE}

#Histogram
library(RColorBrewer)
hist(data$measure,breaks=3 ,col=brewer.pal(3,"Set2"),main="Histogram", xlab="COPD Incidence")

```

# Q-Q Plot 

```{r Q-Q plot, echo=TRUE}
#Q-Q plot 
#install.packages("ggpubr")
library(ggpubr)
ggqqplot(data$measure)
#Q-Q Plot 
qqnorm(data$measure, pch =1, frame = TRUE, lwd=2)
qqline(data$measure, col = "red", lwd = 3)
```

## Normality Test by Shapiro-Wilk normality test 

```{r Shapiro-Wilk normality test}
#Shapiro-Wilk normality test
norm.test <- shapiro.test(data$measure)
norm.test

```

From the output, the p-value < 0.05 implying that the distribution of the data are significantly different from normal distribution. In other words, we can not assume the normality.



## Forest Plot 

```{r forest-plot, echo=TRUE}
label <- as.factor (c(2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012 ,2013, 2014, 2015, 2016, 2017, 2018, 2003, 2004, 2005, 2006, 2007, 2008 ,2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018))
incidence  <- data$measure
lower <- data$lower_confidence_interval
upper <- data$upper_confidence_interval

#Data frame
df <- data.frame(label, incidence, lower,upper)

library(ggplot2)
fp <- ggplot(data=df, aes(x=label, y=incidence, ymin=lower, ymax=upper, col="COPD Incidence")
             ) +
        geom_pointrange() + 
        geom_hline(yintercept=180.3, lty=2, col="blue", lwd=1) +  # add a dotted line at x=1 after flip
        coord_flip() +  # flip coordinates (puts labels on y axis)
        xlab("Year") + ylab("Incidence (95% CI)") +
        theme_bw()  # use a white background
print(fp)
ggsave(path="E:/DG Health Assignment/", "forest plot.png", width = 20, height = 20, units = "cm", dpi=300)
```



## Descriptive Statistical Analysis 

```{r Descriptive statistics}
#Descriptive statistics of Incidence
summary(data$measure)
mean(data$measure)
#Box plot of incidence
boxplot(data$measure,
main = "Incidence of COPD in Scotland"
, xlab = "Incidence  rate per 100,000", ylab = "Incidence",col = "orange", border = "brown", horizontal = FALSE, notch = FALSE)

```

## Linear Model for trend analysis

```{r fit-model}
fit <- lm(data$measure~data$year)
summary(fit)
```


##  Trend Analysis graph 


```{r Trend-analysis, echo=TRUE}
#line plot
p <- ggplot(data = data, aes(x = data$year, y = data$measure)) + 
     geom_line(color = "#00AFBB", size = 1) +
    scale_x_continuous("data$year", labels = as.character(data$year), breaks = data$year)
ggsave(path="E:/DG Health Assignment/", "line plot1.png", width = 30, height = 20, units = "cm", dpi=300)
print(p)

# Polynomial Regression Fitting
#p + stat_smooth(color = "#FC4E07", fill = "#FC4E07",
# method = "loess") + xlab("Year") + ylab("COPD Incidence") 
 #ggsave(path="E:/DG Health Assignment/", "line plot2.png", width = 30, height = 20, units = "cm", dpi=300)

  # Linear Model Fitting
p + stat_smooth(color = "#FC4E07", fill = "#FC4E07",
  method = "lm") + xlab("Year") + ylab("COPD Incidence") 
ggsave(path="E:/DG Health Assignment/", "line plot3.png", width = 30, height = 20, units = "cm", dpi=300)

```
